---
title: "STAT-5361 Homework#7"
author: "Simiao Gao"
output: pdf_document
---

## Exercises 5.3.1

### (1)

$$\begin{aligned}
C \int_0^{\infty} (2x^{\theta-1}+x^{\theta-1/2})e^{-x}dx &= C\int_0^{\infty} 2x^{\theta-1}e^{-x}dx + C\int_0^{\infty} x^{\theta-1/2} e^{-x}dx \\
&= 2C \cdot \Gamma(\theta) + C \cdot \Gamma(\theta+0.5) \\
&= 1 \\
\implies C=\frac{1}{2\Gamma(\theta) + \Gamma(\theta+0.5)}
\end{aligned}$$

$$g(x) = 2C \cdot \Gamma(\theta) \cdot Gamma(\theta,1) + C\cdot \Gamma(\theta+0.5) \cdot Gamma(\theta+0.5,1)$$
Therefore, Component distributions are $Gamma(\theta,1)$ and $\Gamma(\theta+0.5)$, and the coresponding weights are $w_1=\frac{2 \Gamma(\theta)}{2\Gamma(\theta) + \Gamma(\theta+0.5)}$ and $1-w_1$.

### (2)

Take $\theta=3$.

```{r message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=50)}
theta <- 3
theta1 <- theta
theta2 <- theta + 0.5
c <- 1 / (2 * gamma(theta1) + gamma(theta2)) 
w1 <- c * 2 * gamma(theta1)
w2 <- 1 - w1

n <- 10000
set.seed(2020)
u <- rbinom(n, prob = w1, size = 1)
x <- rgamma(n, shape = ifelse(u == 1, theta1, theta2))
g_true <- function(x) {
  return(c * exp(-x) * (2 * x^(theta - 1) + x^(theta - 0.5)))
}
g_sample <- w1 * dgamma(x, shape = theta1) + w2 * dgamma(x, shape = theta2)
plot(x, g_sample, ylab = "density", col = "red", cex = .1)
curve(g_true, from = 0, to = 15, col = "blue", ylab = "", add = TRUE)
legend("topright",
       legend=c("sample","true density"),
       lty=c(1,1),
       col=c("red","blue"))
```

### (3)

$\alpha = sup \frac{f(x)}{g(x)}=sup \frac{\sqrt{4+x} \,x^{\theta-1} e^{-x}}{(2x^{\theta-1}+x^{\theta-1/2})e^{-x}}=sup \frac{\sqrt{4+x}}{2+\sqrt{x}}=1$

```{r message=FALSE, warning=FALSE}
library(ggplot2)
set.seed(2020)
f <- function(x) {
  return(sqrt(4+x) * x^(theta-1) * exp(-x))
}

ratio <- function(x) {
  return(sqrt(4+x) / (2 + sqrt(x)))
}

out <- vector()
for (i in 1:n) {
  while (TRUE) {
    u <- rbinom(1, prob = w1, size = 1)
    x <- rgamma(1, shape = ifelse(u == 1, theta1, theta2))
    uni <- runif(1)
    if (uni < ratio(x)) break
  }
  out[i] <- x
}

d <- unlist(lapply(out, f))
dat <- data.frame(out, d)

cols <- c("sample"="blue","kernel density"="#red")
ggplot(data = dat, aes(x = out)) + 
  geom_density(aes(y=..density.. *5.5), colour = 'red') + 
  geom_line(aes(y = d), colour = 'blue') + 
  xlab("x") + 
  ylab("") 
  
```

Red curve is the scaled kernel density, and blue curve is estimated density from sampling. 

## Exercises 6.3.1

$f(x|\mu_1,\mu_2, \sigma_1^2, \sigma_2^2, \delta ) = \delta \cdot \phi(\mu_1, \sigma_1^2) + (1-\delta) \phi(\mu_2, \sigma_2^2)$

Since the prior distribution for $\delta, \mu_1, \mu_2, \sigma_1^2, \sigma_2^2$ are independent, the posterior distribution 
$f(\boldsymbol{\theta} |x) \propto f(x|\mu_1,\mu_2, \sigma_1^2, \sigma_2^2, \delta ) \, \pi(\delta) \, \pi(\mu_1) \, \pi(\mu_2) \, \pi(\sigma_1^2) \, \pi(\sigma_2^2)$.

```{r message=FALSE, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=50)}
set.seed(2020)
delta <- 0.7 # true value to be estimated based on the data
n <- 100
set.seed(123)
u <- rbinom(n, prob = delta, size = 1)
x <- rnorm(n, ifelse(u == 1, 7, 10), 0.5)

```



